\documentclass[11pt,largemargins, letter]{homework}


\newcommand{\hwname}{Tobias Jakobi}
\newcommand{\hwemail}{tjakobi@arizona.edu}
\newcommand{\hwtype}{Project}
\newcommand{\hwnum}{1}
\newcommand{\hwclass}{CTS505}
\newcommand{\hwlecture}{2023}
\newcommand{\hwsection}{-10}

\usepackage[margin=1in]{geometry}
\usepackage{fontawesome5}



\renewcommand{\questiontype}{Task}

\begin{document}
\maketitle

The solutions for the tasks listed below should be saved in separate \texttt{.R} files, e.g. \texttt{task1.R}, \texttt{task2.R}.

Please include loaded libraries, loading of files, and all other commands in the files.

\question

  \begin{alphaparts}
    \questionpart Download a list of gene IDs with expression values and false discovery rate (FDR) columns from \texttt{https://links.jakobilab.org/cts505\_genes}.
    \questionpart Identify the organism used for the experiment based on the provided gene IDs for the next tasks.
  \end{alphaparts}


\question

  \begin{alphaparts}
    \questionpart Read the provided input file into R with an appropriate function.
    \questionpart Use the \texttt{biomaRt} package to retrieve gene names and Uniprot IDs for the provided gene IDs using the identified organism.
    \questionpart Print the top 10 highest expressed genes based on the logFC value into a new CSV file, \texttt{top10.csv}.
    Include the gene names and Uniprot IDs.
  \end{alphaparts}

\question

  \begin{alphaparts}
    \questionpart Use the \texttt{ggplot2} package to create an X-Y graph that plots \texttt{log10(FDR value)} on the Y axis and the logFC value on the X axis.
    \questionpart Set a title and descriptive X and Y axis labels and save the plot to PDF file named \texttt{xy\_plot.pdf}.
  \end{alphaparts}

\pagebreak
  \question

\begin{itemize}

    \item Download the list of gene counts stored in the following file \\\texttt{https://links.jakobilab.org/p2\_counts}
    \item Identify the organism used for the experiment based on the provided gene IDs for the next tasks.
\end{itemize}

  \begin{alphaparts}

    \questionpart Establish a reasonable grouping of the provided samples based on the sample names.
    \questionpart Read the provided input file into \faRProject \hspace{0.1mm} with an appropriate function.
    \questionpart Use the \texttt{biomaRt} package or any other suitable ID conversion package to retrieve gene names for gene IDs.
  \end{alphaparts}

\question

  \begin{alphaparts}
    \questionpart Develop a suitable design formula as discussed in the class to perform differential gene expression analysis between your chosen groups.
    \questionpart Use the \texttt{edgeR} workflow to build an \texttt{DGElist} object.
    \questionpart Using \texttt{edgeR}, perform filtering and normalization of the imported count data.
    \questionpart Lastly, perform differential gene expression analysis and note the number of up- and down regulated genes.
  \end{alphaparts}

\question

  \begin{alphaparts}
    \questionpart Save a result table with differentially expressed genes filtered by p-value and FDR (0.05) and sorted by logFC.
  \end{alphaparts}

  
  
Please send your R scripts \& plots as attachment via email to \texttt{tjakobi@arizona.edu}.

In case of questions please do not hesitate to contact me.

\end{document}
